---
title: "STATS 485 Paper 3 Appendix Verison 2"
author: "Emmalyn Campau"
date: "2023-12-10"
output:
  pdf_document: default
  word_document: default
---
This file contains code used to perform an analysis that sought to objectively rank NCAA Division 1A football teams using their season records and home field advantage. First, a Bradley-Terry model was constructed using GLM. Then the model was adjusted to include home field advantage. Next, a penalized version of these models was analyzed using Bayesian GLM. All of the models were compared statistically using AIC and 10-fold cross validation. The ranking created by the models with home team advantage were compared to the College Football Playoff committee's ranking. Finally, teams in the Big Ten were investigated and a two-sided adjusted t-test was used to determine the significance of the coefficients created by the B-T models.

# Download libraries
``` {r libraries}
library(dplyr)
library(multcomp)
library(rcompanion)
library(boot)
library(arm)
```

# Download the data
```{r data}
football <- read.csv('https://dept.stat.lsa.umich.edu/~bbh/s485/data/cfb_wl-2023-11-26.csv')
```

# Part A: Create a Bradley-Terry model
a.
``` {r partA}
# set seed
set.seed(123)

# Create a Bradley Terry model
bt_football <- glm(home_win ~ .-1, data = football, family = binomial)
# Extract team strength measures (Michigan State not included, so point out that this was automatically used as the reference)
coef(bt_football)
#sort(coef(bt_football))
```

# Part B: create a B-T model that includes home field advantage
b.
``` {r partB}
# Create a Bradley Terry model
bt_football_adv <- glm(home_win ~ .-Michigan.State, data = football, family = binomial)
# Extract team strength measures
coef(bt_football_adv)

# compare
summary(bt_football)$aic
summary(bt_football_adv)$aic
```

# Create a penalized B-T model using Bayesian GLM
## 10-Fold Cross validation and AIC
c.
``` {r partC}
# Set seed
set.seed(1549)

# Penalized model without home field advantage
bt_bayes <- arm::bayesglm(home_win ~ .-1-Michigan.State, data = football, family = binomial)
coef(bt_bayes)
# Penalized model with home field advantage
bt_bayes_adv <- arm::bayesglm(home_win ~ .-Michigan.State, data = football, family = binomial)
coef(bt_bayes_adv)

# Report statistics, such as AIC
summary(bt_bayes)$aic
summary(bt_bayes_adv)$aic

# Cross validation with 10 folds
# for glm
cv_glm <- cv.glm(football, bt_football, K=10)
(cv_glm$delta)[1]
cv_glm_adv <- cv.glm(football, bt_football_adv, K=10)
(cv_glm_adv$delta)[1]

# for bayesglm
cv_bayes <- cv.glm(football, bt_bayes, K=10)
(cv_bayes$delta)[1]
cv_bayes_adv <- cv.glm(football, bt_bayes_adv, K=10)
(cv_bayes_adv$delta)[1]
```

# Create lists of top 25 teams
d.
``` {r partD}
# Top 25 teams
# GLM with advantage
coef_list_glm <- coef(bt_football_adv)
coef_list_glm <- sort(coef_list_glm, decreasing = TRUE)
# print names
names(coef_list_glm)[1:25]
print("Top 25 Teams Using GLM")
# names and coefficients
head(coef_list_glm, 25)

# Bayes with advantage
coef_list_bayes <- coef(bt_bayes_adv)
coef_list_bayes <- sort(coef_list_bayes, decreasing = TRUE)
# print names
names(coef_list_bayes)[1:25]
print("Top 25 Teams Using BayesGLM")
# names and coefficients
head(coef_list_bayes, 25)

# Print differences
setdiff(names(coef_list_bayes)[1:25], names(coef_list_glm)[1:25])

# Store the CFP's ranking in 'actual'
actual <- c('Georgia', 'Michigan', 'Washington', 'Florida.State', 'Oregon', 'Ohio.State',
            'Texas', 'Alabama', 'Missouri', 'Penn.State', 'Ole.Miss', 'Oklahoma', 'LSU', 
            'Louisville', 'Arizona', 'Iowa', 'Notre.Dame', 'Oklahoma.State', 'NC.State',
            'Oregon.State', 'Tennessee', 'Tulane', 'Clemson', 'Liberty', 'Kansas.State')

# Compare actual to Bayes GLM ranking
setdiff(actual, names(coef_list_bayes)[1:25])
```

# Peer comparison of Big Ten teams
e.
``` {r partE}
# Preferred model uses Bayesian GLM and includes home field advantage. 
# Michigan State as the reference for consistency because I used them above
big_ten <- bayesglm(home_win ~ ., data=football, family=binomial())
# Save coefficients of Big Ten teams
mich <- coef(big_ten)[c('Michigan')]
mich
msu <- coef(big_ten)[c('Michigan.State')]
msu
psu <- coef(big_ten)[c('Penn.State')]
psu
osu <- coef(big_ten)[c('Ohio.State')]
osu

#Bonferroni means that alpha is divided by the number of tests
alpha_bonferroni <- 0.05/3

# Using Michigan as my favorite team. Comparing to OSU, Michigan State, and Penn State using t-test
glht(big_ten, linfct = c("Michigan - Ohio.State == 0")) %>% summary()
# 0.492
glht(big_ten, linfct = c("Michigan - 0 == 0")) %>% summary()
# 0.02
glht(big_ten, linfct = c("Michigan - Penn.State == 0")) %>% summary()
# 0.261

```









